---
title: "Final Project Code Sup"
author: "Elliot Pickens & Andrew Lin & Matt Thill"
date: "June 4, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(Sleuth3)
library(dplyr)
library(ggformula)
library(pander)
library(knitr)
library(stargazer)
library(car)
library(pander)
library(gridExtra)
library(broom)
library(ggthemes)
library(MASS)
library(leaps)
library(GGally)
library(aod)
library(sandwich)
library(pscl)
library(FactoMineR)
library(readxl)
library(survey)
```

## Load the Data 
```{r}
nba <- read_excel("updatedteamdata3.xlsx")
```

##EDA

#Inspect the data

```{r}
summary(nba)
which(is.na(nba))
#no NA values
```


#Let's take a look that the counts that we are trying to model

```{r}
plot(sort(nba$Playoff_Wins))
gf_histogram(~Playoff_Wins, data = nba)
```

##Start by trying to fit a poisson model

```{r}
#Regular Poisson
basic.pois <- glm(Playoff_Wins ~ 1, data = nba, family = poisson)
reg.pois <- glm(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, family = "poisson", data = nba)
summary(reg.pois)
#Goodness of fit test
1 - pchisq(15.371, 16)
#The goodness of fit test doesn't look terrible but let's check anova
anova(reg.pois, test = "Chisq")
#Clearly most of the features are important in this (lazily fit) model
#The regular model doesn't look rediculous, but it does seem a little bit funky (notice the ConfWest signficance, and that nearly
#half of the features appear to be ar least somewhat important)

#Try to fit a better regular poisson model

mod.pois <- stepAIC(basic.pois, scope = c(lower = ~1, upper = ~TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(mod.pois)

#Check fit again
1 - pchisq(165.14, 84)
#The more significant model is a poor fit based in this test
```

##Maybe a quasipoisson model will be better...

```{r}

#Maybe we should try quasipoisson to account for the problems we are having with the regular poisson model

#Quasi Poisson
mod.quasi <- glm(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, family = quasipoisson, data = nba)
summary(mod.quasi)
residualPlots(mod.quasi, type = "deviance", tests = FALSE)
#some of these plots look quite bad (such as Off_Eff)

new.quasi <- glm(Playoff_Wins ~ Off_Eff + Def_Rtg + Conf + TwoP + ThreeP, data = nba, family = quasipoisson)
residualPlots(new.quasi, type = "deviance", tests = FALSE)
#These are better but several of the plots have large (< -2 and > 2) deviance residuals

```

##Before we move one let's see if a negative binomial model can fix our problems

```{r}
#Negative Binomial 

mod.nb <- glm.nb(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba)
summary(mod.nb)
residualPlots(mod.nb, type = "deviance", tests = FALSE)


#Still similar to the previous regular models
#I guess we need to try and fit a different type of model...
```

##Now we will take a look at some models that can hopefully handle the large number of zeros in our data
##The model types that we will investigate are the hurdle and zero inflated model

#start fitting hurdle models using BIC

```{r}
mod.hur <- hurdle(Playoff_Wins ~ FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba, dist = "poisson")
summary(mod.hur)
hur.basic <- hurdle(Playoff_Wins ~ 1, data = nba, dist = "poisson")

hur.step <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(hur.step)

hur.stepNoComp <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf), direction = "both", k = log(nrow(nba)))
summary(hur.stepNoComp)

hur.stepNoCompAIC <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf), direction = "both")
summary(hur.stepNoCompAIC)

hur.stepBk <- stepAIC(mod.hur, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(hur.stepBk)

hur.full <- hurdle(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba, dist = "poisson")
summary(hur.full)
hur.BK <-stepAIC(hur.full, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(hur.BK)

hur.fwd <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "forward", k = log(nrow(nba)))
summary(hur.fwd)

hur.bw <- stepAIC(hur.full, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "backward", k = log(nrow(nba)))
summary(hur.fwd)
```

#Re-Fit the model using AIC

```{r}
hur.AIC <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both")
summary(hur.AIC)

hur.FAIC <- stepAIC(hur.full, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both")
summary(hur.FAIC)
```

#Try negbinomial hurdle model instead of hurdle poisson (BIC)

```{r}
mod.negbin <- hurdle(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba, dist = "negbin")
summary(mod.negbin)
negbin.basic <- hurdle(Playoff_Wins ~ 1, data = nba, dist = "negbin")

negbin.step <- stepAIC(negbin.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(negbin.step)

negbin.full <- hurdle(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba, dist = "negbin")
summary(negbin.full)
negbin.BK <-stepAIC(negbin.full, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(negbin.BK)
```

#Check zeroinfl model (using BIC during stepwise selection)

```{r}
mod.zero <- zeroinfl(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba, dist = "poisson")
summary(mod.zero)
zero.basic <- zeroinfl(Playoff_Wins ~ 1, data = nba, dist = "poisson")

#stepwise selection
zero.step <- stepAIC(zero.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(zero.step)

#the four lines below are error prone 
zero.full <- zeroinfl(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff, data = nba, dist = "poisson")
summary(zero.full)
zero.BK <-stepAIC(zero.full, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff), direction = "both", k = log(nrow(nba)))
summary(zero.BK)
```

#fit testing for: zero.BK, zero.step, hur.BK, hur.step, negbin.step, negbin.BK

```{r}
AIC <- c(AIC(zero.BK), AIC(zero.step), AIC(negbin.step), AIC(negbin.BK), AIC(hur.BK), AIC(hur.step))
plot(AIC)

Loglik <- c(zero.BK$loglik, zero.step$loglik, negbin.step$loglik, negbin.BK$loglik, hur.BK$loglik, hur.step$loglik)
plot(Loglik)
```

#In both of the graphs above the model associated with index 6 looks to be the best (in terms of the Loglik and AIC values)
#The model at index 6 is the hur.step model
#Based on these graphs we are going to move forward with the hur.step model

#Now we will check for interaction terms using repeated stepwise selection to see if any interaction terms are added to the model

```{r}
#Check for interactions

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + Def_Rtg*Conf + PF*Conf + TOV*Conf + BLK*Conf + STL*Conf + AST*Conf + FT*Conf + DRB*Conf + ORB*Conf + ThreeP*Conf + TwoP*Conf), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + Off_Eff*Def_Rtg + Off_Eff*PF + Off_Eff*TOV + Off_Eff*BLK + Off_Eff*STL + Off_Eff*AST + Off_Eff*FT + Off_Eff*DRB + Off_Eff*ORB + Off_Eff*ThreeP + Off_Eff*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + Def_Rtg*PF + Def_Rtg*TOV + Def_Rtg*BLK + Def_Rtg*STL + Def_Rtg*AST + Def_Rtg*FT + Def_Rtg*DRB + Def_Rtg*ORB + Def_Rtg*ThreeP + Def_Rtg*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + PF*TOV + PF*BLK + PF*AST + PF*FT + PF*DRB + PF*ORB + PF*ThreeP + PF*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + TOV*BLK + TOV*STL + TOV*AST + TOV*FT + TOV*ORB + TOV*DRB + TOV*ThreeP + TOV*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + BLK*STL + BLK*AST + BLK*FT + BLK*DRB + BLK*ORB + BLK*ThreeP + BLK*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + STL*AST + STL*FT + STL*DRB + STL*ORB + STL*ThreeP + STL*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + AST*FT + AST*DRB + AST*ORB + AST*ThreeP + AST*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + FT*DRB + FT*ORB + FT*ThreeP + FT*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + DRB*ORB + DRB*ThreeP + DRB*TwoP), direction = "both", k = log(nrow(nba)))

hur.inter <- stepAIC(hur.basic, scope = list(lower = ~ 1, upper = ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff
                                             + Off_Eff*Conf + ORB*ThreeP + ORB*TwoP + ThreeP*TwoP), direction = "both", k = log(nrow(nba)))


```

#summary of interaction model (I removed all of summaries I had inbetween each model re-fitting) 

```{r}
summary(hur.inter)
```

#It looks like Conf*Off_Eff might be significant, so the new model we are using is hur.inter

#Now let's inspect our model further to see if its valid

```{r}
#hur.Finter is a "full" hurdle model that we will use for comparisons
hur.Finter <- hurdle(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff + Off_Eff*Conf, data = nba, dist = "poisson")


#check to see if our model is a better fit than the full and basic models
#NULL Hypothesis: more complex model is a better fit
pchisq(2*(hur.inter$loglik - hur.basic$loglik), df = 16, lower.tail = FALSE)
#based on the low p val it looks like our model is definetly better than the basic model
pchisq(2*(hur.Finter$loglik - hur.inter$loglik), df = 16, lower.tail = FALSE)
#based on the high p val it looks like our model is a better fit than the full model
```

#Our model isn't looking too bad, but lets use a puesdo cross validation approach to see if our model
#is more significant than the "full" model when we look at a random 9/10ths of the data 

```{r}
numSim <- 1000
result <- numeric(numSim)
for (i in 1:numSim) {
  tempData <- nba[sample(nrow(nba), size = 81),]
  mod.red <- hurdle(Playoff_Wins ~ Off_Eff + Def_Rtg +Conf + STL + TOV + Off_Eff*Conf, data = tempData, dist = "poisson")
  mod.f <- hurdle(Playoff_Wins ~ TwoP + ThreeP + ORB + DRB + FT + AST + STL + BLK + TOV + PF + Conf + Def_Rtg + Off_Eff + Off_Eff*Conf, data = tempData, dist = "poisson") 
  pVal <- pchisq(2*(mod.f$loglik - mod.red$loglik), df = 16, lower.tail = FALSE)
  result[i] <- pVal
  rm(tempData)
}
summary(result)
gf_histogram(~result)
```

#Now we are going to use the puesdo cross validation approach to get estimates for our coefficients

```{r}
##COUNT CI bands:
resultslower <- numeric(7)
resultsupper <- numeric(7)
for(k in 1:7){

  N <- 1000
  result <- numeric(N)
  for(i in 1:N){
    tempdata <- nba[sample(nrow(nba),size=81),]
    testhurdle <- hurdle(Playoff_Wins~Off_Eff+Def_Rtg+Conf+STL+TOV+Conf*Off_Eff,data=tempdata,dist="poisson")
    tempout <- summary(testhurdle)
    result[i] <- tempout$coefficients$count[k,1]
    
    
  }
  result <- sort(result)
  
  resultslower[k] <- result[25]
  resultsupper[k] <- result[975]

}
resultslower
resultsupper



##Zero CI Bands:

Zeroresultslower <- numeric(7)
Zeroresultsupper <- numeric(7)
for(k in 1:7){

  N <- 1000
  Zeroresult <- numeric(N)
  for(i in 1:N){
    tempdata <- nba[sample(nrow(nba),size=81),]
    testhurdle <- hurdle(Playoff_Wins~Off_Eff+Def_Rtg+Conf+STL+TOV+Conf*Off_Eff,data=tempdata,dist="poisson")
    tempout <- summary(testhurdle)
    Zeroresult[i] <- tempout$coefficients$zero[k,1]
    
    
  }
  Zeroresult <- sort(Zeroresult)
  
  Zeroresultslower[k] <- Zeroresult[25]
  Zeroresultsupper[k] <- Zeroresult[975]

}
Zeroresultslower
Zeroresultsupper
```


#Re-do the AIC and loglik graphs with the interaction model

```{r}
AIC2 <- c(AIC(zero.BK), AIC(zero.step), AIC(negbin.step), AIC(negbin.BK), AIC(hur.BK), AIC(hur.step), AIC(hur.inter))
plot(AIC2)

Loglik2 <- c(zero.BK$loglik, zero.step$loglik, negbin.step$loglik, negbin.BK$loglik, hur.BK$loglik, hur.step$loglik, hur.inter$loglik)
plot(Loglik2)
```

#In the graphs above the interaction model is the model at index 7
#Based on the graphs it looks like the interaction model wins out on the loglik and AIC fronts






















